Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2014-10-01 exp1/present2013.tex
bump date to the Vienna talk
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@762 43ef1190-90fd-4dd7-a769-d0121f02b80e
reintroduce details on the experiment
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@763 43ef1190-90fd-4dd7-a769-d0121f02b80e
bugfix
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@764 43ef1190-90fd-4dd7-a769-d0121f02b80e

\title[DoE in SPARQL Evaluations]{Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation}
\subtitle{}
\author{Kjetil Kjernsmo and John Tyssedal}
\date{24.10.2013}
\maketitle

@@

\begin{frame}{Factors and levels}
  \begin{description}
  \item [``Implement''] The implementation under evaluation.
  \item [``TripleC''] 1 or 2 MTriples from the DBPedia SPARQL
    Benchmark.
  \item [``Machine''] Software and hardware platform, one smaller
    machine with fast SSD and one larger with slower disks in RAID1.
  \item [``BGPComp''] Two different basic graph patterns with varying
    complexity.
  \item [``Union''] Absence or presence of a \textsf{UNION} pattern.
  \item [``Lang''] Absence or presence of a \textsf{langMatches}
    filter.
  \item [``Range''] Absence or presence of a filter with a larger-than
    operator.
  \item [``Optional''] Absence or presence of an \textsf{OPTIONAL}
    pattern.

@@

\note{
  \begin{itemize}
  \item Those that are deemed significant are labelled.
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, 
  \item Interactions with Implement with joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
%  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
%    to that many more triples need to be searched for a language tag
%    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''? Leave that hanging for
    a bit.
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine'', in the paper we describe how this
    enables us to average over certain factors and look upon the rest
    as a replicated experiment.
  \end{itemize}
}

\begin{frame}{Hypothesis tests}

@@

% TODO: Much around the hypothesis test

Tests to run:
\begin{itemize}

>>>>

\title[DoE in SPARQL Evaluations]{Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation}
\subtitle{}
\author{Kjetil Kjernsmo and John Tyssedal}
\date{07.10.2014}
\maketitle

@@

\begin{frame}{Factors and levels}
  \begin{description}
  \item [``Implement''] The implementation under evaluation.
  \item [``TripleC''] 1 or 2 MTriples from the DBPedia SPARQL
    Benchmark.
  \item [``Machine''] Software and hardware platform, one smaller
    machine with fast SSD and one larger with slower disks in RAID1.
  \end{description}
\end{frame}

\begin{frame}[fragile]{The Basic Graph pattern factor}
  \begin{description}

  \item [``BGPComp''] Two different basic graph patterns with varying
    complexity.
  \end{description}

  \begin{center}\small
    \begin{tabular}{|@{~~}p{0.40\textwidth}|@{~~}p{0.40\textwidth}|}
      \hline
      \textbf{Level 1} & \textbf{Level 2} \\[-2ex] 
      \begin{verbatim}
?s  rdfs:label ?l1 ;
    ?p1 ?o1 .
?o1 dbo:populationTotal ?l2 .
      \end{verbatim}
      & 
      \begin{verbatim}
?s  rdfs:label ?l1 ;
    ?p1 ?o1 .
?o1 dbo:populationTotal ?l2 .
?s foaf:page ?o2 ;
   dbpprop:subdivisionName ?o3 .
?o3 skos:subject ?o4 ;
   dbpprop:seat ?o5 ;
   a ?c1 .
\end{verbatim} 
\\      \hline
    \end{tabular}
  \end{center}

\begin{frame}[fragile]{Absence or presence factors}
  \begin{description}
  \item [``Union''] Absence or presence of a \textsf{UNION} pattern.

    \begin{verbatim}
{ ?o1 dbpprop:longd ?long ;
      dbpprop:latd  ?lat .
} UNION {
  ?o1 geo:long ?long ;
      geo:lat ?lat . }
\end{verbatim}

  \item [``Lang''] Absence or presence of a \textsf{langMatches}
    filter.

\begin{verbatim}
FILTER langMatches( lang(?l1), "fr" ) 
\end{verbatim}

  \item [``Range''] Absence or presence of a filter with a larger-than
    operator.

    \begin{verbatim}
FILTER (?l2 > 800000)
\end{verbatim}

  \item [``Optional''] Absence or presence of an \textsf{OPTIONAL}
    pattern.
    \begin{verbatim}
OPTIONAL { ?o1 foaf:homepage ?o6 . }
\end{verbatim}

@@

\note{
  \begin{itemize}
% TODO: Lenth
  \item Those that are deemed significant are labelled.
  \item Don't worry about the axes
  \item Yes, the Implementation is better, we knew that, 
  \item Interactions with Implement with joins are
    naturally strong
  \item ``Union'', ``Optional'' and ``Lang'' demanding things
%  \item ``BGPComp:Lang'' and ``Lang:Union'' interactions might be due
%    to that many more triples need to be searched for a language tag
%    in one of the levels of the interacting factors.
  \item But what about ``Implement::TripleC''? Leave that hanging for
    a bit.
  \item We want to do a hypothesis test
  \item Two factors do not contribute significantly to the variation
    ``Range'' and ``Machine'', in the paper we describe how this
    enables us to average over certain factors and look upon the rest
    as a replicated experiment.
  \end{itemize}
}

\begin{frame}{Preparing hypothesis tests}

@@

Classifying factors:
  \begin{itemize}
  \item ``Range'' and ``Machine'' are called \emph{inactive factors}.
  \item ``BGPComp'', ``Lang'', ``Optional'', ``Union'' and ``TripleC''
    are called \emph{environmental factors}.
  \item ``Implement'' is called a \emph{control factor}.
  \end{itemize}

\end{frame}
\begin{frame}{Hypothesis tests}

\begin{itemize}
\item Average over all the 32 level-combinations of the environmental
  factors. 
\end{itemize}

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Tue Oct  8 14:51:28 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rlll|r}
  \hline
 & Implement & Machine & Range & Response \\ 
  \hline
1 & 1 & 1 & 1 & 33.47 \\ 
  2 & 2 & 1 & 1 & 11.82 \\ 
  3 & 1 & 2 & 1 & 32.16 \\ 
  4 & 2 & 2 & 1 & 11.56 \\ 
  5 & 1 & 1 & 2 & 33.97 \\ 
  6 & 2 & 1 & 2 & 12.46 \\ 
  7 & 1 & 2 & 2 & 32.90 \\ 
  8 & 2 & 2 & 2 & 11.60 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

\begin{itemize}
